---
title: "MCMC run"
author: "Eduardo de Freitas Costa"
date: '`r format(Sys.Date(), "%Y-%m-%d")`'
output:
  html_document:
    df_print: paged
    code_folding: hide
    toc: yes
    toc_depth: 3
    toc_float: true
editor_options:
  chunk_output_type: console
---

# Script header

File: `Descriptive.rmd`

Client: 
WBVR project: 
Author: Eduardo de Freitas Costa, Wageningen Bioveterinary Research

Start date: \
R version: `r getRversion()`

Dependencies:

-   Downstream

    -   `Data\hiuwalter_model_insilico.txt`

-   Upstream

    -   `None`

-   Input

        -   `None`.

-   Output

    -   `None`

Peer reviewer(s)

------------------------------------------------------------------------

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```


# Load libraries
```{r, include=FALSE}

#Packages to be used
packages<-c("here","tidyverse","ggplot2","knitr","rmarkdown","runjags")

# Install packages not yet installed
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}

# Packages loading
invisible(lapply(packages, library, character.only = TRUE))


```

# Import the data
```{r}

```

# Run the MCMC model

```{r}

if (file.exists(here("Data","huiwalter_model.txt"))){
results <- run.jags(here("Data","huiwalter_model.txt"),
                    burnin = 1500,
                    sample = 10000,
                    n.chains = 2,
                    thin = 2)
  
} else{
template_huiwalter(simdata,outfile=here("Data","huiwalter_model.txt"), covariance=FALSE,
                   se_priors = "dbeta(1,1)",
                    sp_priors = "dbeta(1,1)")

results <- run.jags(here("Data","huiwalter_model.txt"),
                    burnin = 1500,
                    sample = 10000,
                    n.chains = 2,
                    thin = 2)
  
  
}

```


# Model results and diagnostic
```{r}
## Examine trace plots to assess model convergence
plot(results, vars = c("se", "sp", "prev"))

## Model gives results for Se, Sp and the PPP values
## Estimated sample size (ESS) and potential scale reduction factor (PSRF) of the Gelman Rubin statistic included in the output
results

```

# PCR1 cut-off
## Extract the PPP values and match these to PCR1 results
```{r}

## PPP values are calculated in JAGS, so need to be extracted from 10,000 (thinned) iterations:
ppp_values <- combine.mcmc(res1, return.samples=10000, vars="ppp")

## Mean PPP values (n = 28) based on 1000 iterations
(mean_ppp <- apply(ppp_values, 2,  mean))

## Map the 28 mean PPVs to all the observations (n = 233) in the df1 dataset
test_data %>%
  mutate(	testcombo = interaction(MAT, WB), 
  			colname = str_c("ppp[", as.numeric(testcombo), ",", as.numeric(Population), "]"), 
  			colindex = match(colname, names(mean_ppp)), 
  			mean_ppp = mean_ppp[colindex]
  		) ->
  test_data

```

## Derive PCR1 Se and Sp for each possible PP value from these 1000 iterations:
```{r}
possible_pp <- tibble(LT = c(0, sort(unique(test_data$ELISA_PP)))) %>%
  mutate(UT = lead(LT, default=max(test_data$ELISA_PP)+1), MT = (LT+UT)/2, IT = 1:n()) 

cutoff_fun <- function(iter){
  test_data %>%
    select(ELISA_PP, colindex) %>%
    mutate(PPP = ppp_values[iter,colindex]) %>%
    expand_grid(possible_pp) %>%
    group_by(IT, LT, MT, UT) %>%
    summarise(	Iteration = iter, 
    			SeNum = sum(PPP[ELISA_PP >= MT]), SeDen = sum(PPP),
    			SpNum = sum(1-PPP[ELISA_PP < MT]), SpDen = sum(1-PPP), 
    			.groups='drop')
}

cutoff_evaluation <- pblapply(seq_len(coda::niter(ppp_values)), cutoff_fun) %>%
	bind_rows() %>%
	mutate(
		Sensitivity = SeNum / SeDen,
		Specificity = SpNum / SpDen,
		SeCI = rbeta(n(), SeNum+1, SeDen-SeNum+1),
		SpCI = rbeta(n(), SpNum+1, SpDen-SpNum+1)
	 )


## Create a plot of sensitivity, specificity and sum vs PP value:
	 
getci <- function(x, upper=TRUE) coda::HPDinterval(coda::as.mcmc(x))[,as.numeric(upper)+1]
plotdata <- cutoff_evaluation %>%
	 group_by(IT, LT, MT, UT) %>%
	 summarise(
	 	Se_Mean = mean(SeNum / SeDen),
	 	Se_LCI = getci(SeCI, FALSE),
	 	Se_UCI = getci(SeCI, TRUE),
	 	Sp_Mean = mean(SpNum / SpDen),
	 	Sp_LCI = getci(SpCI, FALSE),
	 	Sp_UCI = getci(SpCI, TRUE),
	 	.groups = "drop"
	 ) %>%
	 mutate(
	 	Sum_Mean = Se_Mean + Sp_Mean,
	 	Sum_LCI = Se_LCI + Sp_LCI,
	 	Sum_UCI = Se_UCI + Sp_UCI
	 ) %>%
	 pivot_longer(
	 	cols = Se_Mean:Sum_UCI,
	 	names_to = c("Parameter", "Estimate"),
	 	names_pattern = "(.*)_(.*)",
	 	values_to = "Value"
	 ) %>%
	 mutate(Value = Value * 100) %>%
	 pivot_wider(
	 	names_from = Estimate, values_from = Value
	 ) %>%
	 pivot_longer(
	 	cols = LT:UT,
	 	names_to = "Boundary",
	 	values_to = "PP"
	 ) %>%
	 mutate(Parameter = factor(Parameter, levels=c("Sum", "Se", "Sp"), labels = c("Sensitivity + Specificity", "Sensitivity", "Specificity")))

optimal <- plotdata %>%
	filter(Parameter == "Sensitivity + Specificity", Boundary == "MT") %>%
	arrange(desc(Mean)) %>%
	slice(1) %>%
	pull(PP)
optimal
round(optimal, 1)
# Beware floating point arithmetic!
print(optimal, digits=20)
(optimal <- round(optimal + 1e-6, 1))

ggplot(plotdata) +
	aes(x = PP, y = Mean, ymin = LCI, ymax = UCI) +
	geom_ribbon(alpha=0.25) +
	geom_line() +
	geom_vline(xintercept = 20, lty="dotted") +
	geom_vline(xintercept = optimal, lty="dashed") +
	facet_wrap( ~ Parameter, scales="free_y", ncol=1) +
	xlab("PP value") +
	ylab(NULL) +
	scale_x_continuous(breaks = c(0, 20, optimal, seq(50,150,by=25)))


## Create a ROC curve:
roc_ci <- cutoff_evaluation %>%
	mutate(Se = as.integer(round(SeCI*100)), Sp = as.integer(round(SpCI*100))) %>%
	group_by(Sp) %>%
	summarise(LCI = getci(Se, FALSE), UCI = getci(Se, TRUE), Se = mean(Se)) %>%
	arrange(Se, 1-Sp)
roc_mean <- cutoff_evaluation %>%
	group_by(IT,LT,MT,UT) %>%
	summarise(Se = mean(SeNum/SeDen)*100, Sp = mean(SpNum/SpDen)*100, .groups='drop')
roc_opt <- plotdata %>%
	filter(abs(round(PP + 1e-6, 1) - optimal) < 1e-6, Parameter %in% c("Sensitivity","Specificity")) %>%
	select(Parameter, Mean) %>%
	pivot_wider(names_from = Parameter, values_from = Mean)

ggplot() + 
	geom_ribbon(aes(x=100-Sp, y=Se, ymin=LCI, ymax=UCI), roc_ci, alpha = 0.25) +
	geom_line(aes(x=100-Sp, y=Se), roc_mean) +
	geom_point(aes(x=100-Specificity, y=Sensitivity), roc_opt, pch = 8, col="black") +
	ylab('Sensitivity') + xlab('1 - Specificity') +
	geom_abline(slope=1, intercept=0, lty='dashed') +
	ylim(0,100)

# Calculate AUC:
integrate(approxfun(1-roc_mean$Sp/100, roc_mean$Se/100), 0, 1)

```

# PCR2 cut-off
## Extract the PPP values and match these to PCR2 results
```{r}

```
